Title here

### 2020-08-26 paper/paper.Rmd
minor fixes sections 2 and 3

Given an optimisation problem, two basic approaches find the optimum based on different thinking.
An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the close-form solution may not be available when the problem starts to become complex. 
An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution. An iterative method may end up finding a local optimum but the progressive nature of the algorithm allows the practitioner to decide when to stop if a desirable accuracy has been achieved.

A traditional while often used in practice is an iteartive method called *line search method* [@fletcher2013practical]. In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.

@@

Before introducing the guided tour, we shall be familiar with the general tour method [@cook2008grand]. A tour produces animated visualisation of the high dimensional data via rotating low dimension planes. The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to @buja2005computational for the mathematical details. Iteratively choose different low dimension planes and interpolate between them forms a tour path. Different types of tour methods choose the low dimensional planes differently and we mention two other type of tour that are commonly used. A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data. Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.

@@

where $I_d$ is the d-dimentional identity matrix. The constraint requires the projection basis $\mathbf{A}$ to be a orthonogal matrix with each column vector being orthornormal.

There are several features of this optimisation that are worth noticing. First of all, it is a multivariate constraint optimisation problem. Since the decision variables are the entries of a projection basis, it is required to be orthonormal. It is also likely that the objective function is non-differentiable or the gradient information is simply not available. In this case, we will need to either use some approximation of the gradient or turn to derivative free methods. Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem need to be able to find the global maximum. Along the way, locla maximum may also be of our interest since they could present unexpected interesting projections. There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.

@@

Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient acsent.

@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$. The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path. A new round of search continues to find a better basis after the interpolation finishes. The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$. When the search needs to sample more than $c$ number of basis to find an accpeted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration. If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops. The algorithm is summarised in Algorithm \ref{random-search} for one iteration. [mention orthonormalise to ensure the constaint is fulfilled; don't use derivative information but a random search]

@@

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] modifies `search_better` based on a non-increasing cooling scheme $T(j)$. Given an initial $T_0$, the temperature at iteration $k$ is defined as $T(j) = \frac{T_0}{\log(j + 1)}$. When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probabiliby $$P= \min\left\{\exp\left[-\frac{I_{cur} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis. This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space. This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e. a one dimenional projection of the dataset with two informative variables). The algorithm can be written as replacing line 5-11 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.

@@

@cook1995grand used a gradient ascent algorithm with pseudo-derivative. Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the graident. The $\delta$ parameter is usually set to be tiny to ensure the geogesic direction is a good local approximation to the graident. Once the searching direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic as the candidate. $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates. Algorithm \ref{search-geodesic} summarise the steps in geodesic search.

@@

A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum. One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search. 
When plotting the searching observations as points in a plot, the space each iteartion take will be proportional to the number of points in that iteration.

>>>>

Given an optimisation problem, two basic approaches find the optimum based on different thinking.
An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the closed-form solution may not be available when the problem starts to become complex. 
An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution. An iterative method may end up finding a local optimum but the progressive nature of the algorithm allows the practitioner to decide when to stop if a desirable accuracy has been achieved.

A traditional while often used in practice is an iterative method called *line search method* [@fletcher2013practical]. In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.

@@

Before introducing the guided tour, we shall be familiar with the general tour method [@cook2008grand]. A tour produces animated visualisation of the high dimensional data via rotating low dimension planes. The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to @buja2005computational for the mathematical details. Iteratively choosing different low dimension planes and interpolating between them forms a tour path. Different types of tour methods choose the low dimensional planes differently and we mention two other types of tour that are commonly used. A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data. Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.

@@

where $I_d$ is the d-dimensional identity matrix. The constraint requires the projection basis $\mathbf{A}$ to be an orthogonal matrix with each column vector being orthonormal.

There are several features of this optimisation that are worth noticing. First of all, it is a multivariate constraint optimisation problem. Since the decision variables are the entries of a projection basis, it is required to be orthonormal. It is also likely that the objective function is non-differentiable or the gradient information is simply not available. In this case, we will need to either use some approximation of the gradient or turn to derivative free methods. Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem needs to be able to find the global maximum. Along the way, local maximum may also be of our interest since they could present unexpected interesting projections. There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.

@@

Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient ascent.

@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$. The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path. A new round of search continues to find a better basis after the interpolation finishes. The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$. When the search needs to sample more than $c$ number of basis to find an accepted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration. If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops. The algorithm is summarised in Algorithm \ref{random-search} for one iteration. [mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]

@@

Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] modifies `search_better` based on a non-increasing cooling scheme $T(j)$. Given an initial $T_0$, the temperature at iteration $k$ is defined as $T(j) = \frac{T_0}{\log(j + 1)}$. When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis. This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space. This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e. a one dimensional projection of the dataset with two informative variables). The algorithm can be written as replacing line 5-11 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.

@@

@cook1995grand used a gradient ascent algorithm with pseudo-derivative. Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the gradient. The $\delta$ parameter is usually set to be tiny to ensure the geodesic direction is a good local approximation to the gradient. Once the searching direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic as the candidate. $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates. Algorithm \ref{search-geodesic} summarises the steps in geodesic search.

@@

A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum. One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search. 
When plotting the searching observations as points in a plot, the space each iteration take will be proportional to the number of points in that iteration.
